---
title: Gridded OBIS data in R
date: "`r Sys.Date()`"
author: Pieter Provoost
output: (function(...) {
  rmdformats::robobook(toc_depth = 4, fig_caption = FALSE, pandoc_args = c("+RTS", "-K2000m", "-RTS"), ...) })
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })  
---

The robis R package currently does not provide access to gridded OBIS data, but fetching gridded data from the API is straightforward. In this short notebook we will fetch gridded data for a single species and visualize it using Leaflet and ggplot2.

First load some required packages.

```{r message=FALSE, warning=FALSE}
library(httr)
library(dplyr)
library(leaflet)
library(geojsonsf)
library(ggplot2)
library(rnaturalearth)
library(sf)
library(glue)
library(viridis)
```

The API endpoint for gridded data looks like this: `https://api.obis.org/occurrence/grid/{res}` where `res` is the Geohash resolution. Construct a URL, fetch the data, and convert to sf:

```{r message=FALSE, warning=FALSE}
res <- 3
taxon <- "Abra alba"

url <- glue("https://api.obis.org/occurrence/grid/{res}?scientificname={taxon}")
json <- GET(URLencode(url)) %>% content(as = "text")
sf <- geojson_sf(json)
```

This creates an Leaflet based web map:

```{r message=FALSE, warning=FALSE}
leaflet(sf) %>%
  addProviderTiles("Esri.OceanBasemap") %>%
  addPolygons(weight = 1, color = "#cc3300")
```

And here's how to visualize the sf object using ggplot2:

```{r message=FALSE, warning=FALSE}
world <- ne_countries(scale = "medium", returnclass = "sf")

ggplot() + 
  geom_sf(data = world, size = 0) +
  geom_sf(data = sf, aes(fill = n, size = NA)) +
  coord_sf(crs = st_crs("ESRI:54030")) +
  scale_fill_viridis(trans = "log10") +
  theme_void()
```  
